---
title: "Getting Started"
output: html_document
vignette: >
  %\VignetteIndexEntry{Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ddplot)
```

`D3.js` is a famous JavaScript library that allows one to create extremely flexible SVG graphics however `D3` has (at least according to me) a pretty steep learning curve. Further, in order to understand some core concepts, one need to have some basics in `HTML`, `CSS` and `JavaScript`. `ddplot` aims to simply the process using a set of functions that render several graphics using a simple `R` API. Finally, `ddplot` is built upon the amazing `r2d3` package which makes it a breeze to interface `D3.js` with `R`, so a big thanks to the developers.

# `scatterPlot()`

Let's work with the `mpg` data frame from the `ggplot2` package.

```{r fig.align='center', message=FALSE, warning=FALSE}
library(ggplot2) # needed for the mpg data frame

scatterPlot(
  data = mpg,
  x = "hwy",
  y = "cty",
  xtitle = "hwy variable",
  ytitle = "cty variable",
  title = "cty and hwy relationship",
  titleFontSize = 20
)

```

In comparison to `ggplot2`, graphics' customization in `ddplot` is limited nonetheless you get a fully vectorized SVG which is cool.



```{r, fig.align='center'}
scatterPlot(
  data = mpg,
  x = "displ",
  y = "cty",
  col = "tomato",
  bgcol = "pink",
  size = 3,
  stroke = "royalblue",
  strokeWidth = 1,
  xtitle = "displ variable",
  ytitle = "cty variable",
  xticks = 3,
  yticks = 3)

```

# `histogram()`

The `histogram()` function allows you to visualize the distribution of a vector of data:

```{r}
histogram(
  x = mpg$hwy,
  bins = 20,
  fill = "crimson",
  stroke = "white",
  strokeWidth = 1,
  title = "Distribution of the hwy variable",
  width = "20",
  height = "10"
)
```


# `animatedHistogram()`

This function allows you to create a one-click histogram animation. Useful for presentation purposes. Click on the following empty plot and see what happens:

```{r}
animatedHistogram(
  x = mpg$hwy,
  duration = 2000,
  delay = 100,
  fill = "lime",
  stroke = "white",
  bgcol = "white"
  )
```

Note that you can customize the animation using the two parameters `duration` and `delay`.

# `emoji_histogram()`

You can use __emojis__ to render your histogram (try to hover over the emojis):

```{r}
emoji_histogram(
  x = mpg$hwy,
  bins = 20,
  emoji = "🌸",
  emojiSize = 30,
  stroke = "white",
  bgcol = "white", 
  title = NULL
)
```



# `flower()` 

The `flower()` function allows you to plot a simple flowers with a defined number of petals. You can also define the width/length of the petals and rotate them if needed: 


```{r}
flower(
  petalCount = 7,
  petalColor = "plum",
  centerText = "🌼",
  rotationSpeed = 0
)
```


```{r}
flower(
  petalCount = 15,
  petalWidth = 20,
  petalColor = "plum",
  rotationSpeed = 1.5
)
```


# `barChart()`

The `barChat()` function allows you to create bar charts however you need to make the aggregation beforehand. In the following example, we will plot the average `cty` for each `manufacturer` using the `dplyr` package.

```{r message=FALSE, warning=FALSE}
library(dplyr)

mpg %>% group_by(manufacturer) %>%
  summarise(mean_cty = mean(cty)) %>%
  barChart(
    x = "manufacturer",
    y = "mean_cty",
    xFontSize = 10,
    yFontSize = 10,
    fill = "orange",
    strokeWidth = 2,
    ytitle = "average cty value",
    title = "Average City Miles per Gallon by manufacturer"
  )
```

The bars can be easily sorted in `ascending` or `descending` order using the `sort` parameter:


```{r message=FALSE, warning=FALSE}
mpg %>% group_by(manufacturer) %>%
  summarise(mean_cty = mean(cty)) %>%
  barChart(
    x = "manufacturer",
    y = "mean_cty",
    sort = "ascending",
    xFontSize = 10,
    yFontSize = 10,
    fill = "orange",
    strokeWidth = 1,
    ytitle = "average cty value",
    title = "Average City Miles per Gallon by manufacturer",
    titleFontSize = 16
  )
```


# `horzBarChart()`

If you've many categories, it might be a good idea to go for a horizontal bar chart. It has the same parameters as the `barChart()` function except that the x-axis parameter is named `value` and the y-axis parameter named `label`, this naming convention aims to mitigate some confusion that can arise.

If we want to replicate the above graphic in a horizontal way, we can do:

```{r}
mpg %>% group_by(manufacturer) %>%
  summarise(mean_cty = mean(cty)) %>%
  horzBarChart(
    label = "manufacturer",
    value = "mean_cty",
    sort = "ascending",
    labelFontSize  = 10,
    valueFontSize = 10,
    fill = "orange",
    stroke = "crimson",
    strokeWidth = 1,
    valueTitle  = "average cty value",
    title = "Average City Miles per Gallon by manufacturer",
    titleFontSize = 16
  )
```

As in `barChart()`, we can aslo sort in descending order:


```{r}
mpg %>% group_by(manufacturer) %>%
  summarise(mean_cty = mean(cty)) %>%
  horzBarChart(
    label = "manufacturer",
    value = "mean_cty",
    sort = "descending",
    labelFontSize  = 10,
    valueFontSize = 10,
    bgcol = "black",
    axisCol = "white",
    fill = "white",
    stroke = "white",
    strokeWidth = 1,
    valueTitle  = "average cty value",
    labelTitle = "Manufacturers",
    title = "Average City Miles per Gallon by manufacturer",
    titleFontSize = 16
  )
```


# `lollipopChart()`

lollipop chart follows the same behavior as bar charts but instead of bars you get lollipops, hence the name. Below an example of a lollipop chart with `ddplot`:


```{r}
mpg %>% group_by(drv) %>%
  summarise(median_cty = median(cty)) %>%
  lollipopChart(
    x = "drv",
    y = "median_cty",
    sort = "ascending",
    xtitle = "drv variable",
    ytitle = "median cty",
    title = "Median cty per drv",
    xFontSize = 20
  )
```


It's possible to grasp the distribution of some variable according to a specific categorical variable using the same function:


```{r}

mpg %>% filter(year == 2008) %>%
lollipopChart(
    x = "manufacturer",
    y = "hwy",
    circleFill = 'red',
    circleStroke = 'orange',
    circleRadius = 5,
    sort = "none",
    xFontSize = 10
  )
```

From above, it's quite easy to notice that although Toyota has two cars with high highway miles per galon (hwy), it also produces many other vehicles with poor hwy.


# `horzLollipop()`

Same with bar charts, if you have a variable that has many categorical values, you can work with the reversed version of `lollipopChart()` which is `horzLollipop()`:


```{r}
mpg %>% group_by(manufacturer) %>%
  summarise(median_cty = median(cty)) %>%
  horzLollipop(
    label = "manufacturer",
    value = "median_cty",
    sort = "descending")
```
You can also do:



```{r}
mpg %>% filter(year == 2008) %>%
horzLollipop(
    label = "manufacturer",
    value = "hwy",
    circleFill = 'red',
    circleStroke = 'orange',
    circleRadius = 5,
    sort = "none"
  )
```


# `pieChart()`

Pie charts and donut charts are pretty straightforward to set up. We'll use a sample from the `starwars` data frame to plot a simple pie chart.


```{r}
# starwars is part of the dplyr data frame
mini_starwars <- starwars %>% tidyr::drop_na(mass) %>%
  sample_n(size = 5) # getting 5 random values

pieChart(
  data = mini_starwars,
  value = "mass",
  label = "name"
)
```

Using the `padRadius`, `padAngle` and `cornerRadius` parameters, one can get fanciers pie charts:

```{r}
pieChart(
  data = mini_starwars,
  value = "mass",
  label = "name",
  padRadius = 200,
  padAngle = 0.1,
  cornerRadius = 50,
  innerRadius = 10
)
```
If you need a donut chart, you just need to play with the `innerRadius` parameter:

```{r}
pieChart(
  data = mini_starwars,
  value = "mass",
  label = "name",
  innerRadius = 120,
  cornerRadius = 20,
  title = "5 Starwars characters ranked by their mass",
  titleFontSize = 16,
  bgcol = "yellow"
)
```

# `lineChart()`

The `lineChart()` function is used to plot time series data. The use must provide a `date` variable that has the `yyyy-mm-dd` format. In the following example, we'll use the `Air Passenger` built-in `ts` data and convert it to a classical data frame:


```{r}
# 1. converting AirPassengers to a tidy data frame
airpassengers <- data.frame(
  passengers = as.matrix(AirPassengers),
  date= zoo::as.Date(time(AirPassengers))
)

# 2. plotting the line chart
lineChart(
  data = airpassengers,
  x = "date",
  y = "passengers"
)
```

You can modify the line interpolation using the `curve` parameter:

```{r}
lineChart(
  data = airpassengers,
  x = "date",
  y = "passengers",
  curve = "curveStep"
)
```


```{r}
lineChart(
  data = airpassengers,
  x = "date",
  y = "passengers",
  curve = "curveCardinal"
)
```

```{r}
lineChart(
  data = airpassengers,
  x = "date",
  y = "passengers",
  curve = "curveBasis"
)
```

# `animLineChart()`

Heavily inspired from [Jure Stabuc's example](https://observablehq.com/@jurestabuc/animated-line-chart), the `animLineChart()` function create an empty SVG but when each time you click on it a line chart animation starts. Note that the line lasts after the end of the animation. Go ahead, click on the empty graphic below:

```{r}
animLineChart(
  data = airpassengers,
  x = "date",
  y = "passengers",
  duration = 10000, # in milliseconds (10 seconds)
  curve = "curveCardinal"
  )
```


# `areaChart()`

`areaChart()` works similarly except that instead of a line you get an area.


```{r}
# 1. converting AirPassengers to a tidy data frame
airpassengers <- data.frame(
  passengers = as.matrix(AirPassengers),
  date= zoo::as.Date(time(AirPassengers))
)

# 2. plotting the area chart
areaChart(
  data = airpassengers,
  x = "date",
  y = "passengers",
  fill = "purple",
  bgcol = "white"
)
```

# `areaBand()`

`areaBand()` lets you plot a filled area between two y-values. For the sake of the example, let's create an additional column `passengers_upper` that has an additional 40 passengers for each observation:

```{r}
airpassengers <- data.frame(
  passengers_lower = as.matrix(AirPassengers),
  passengers_upper = as.matrix(AirPassengers) + 40,
  date= zoo::as.Date(time(AirPassengers))
)

areaBand(
  data = airpassengers,
  x = "date",
  yLower = "passengers_lower",
  yUpper = "passengers_upper",
  fill = "yellow",
  stroke = "black"
)
```


# `stackedAreaChart()`

This function allows you to create a stacked area chart. You need two components:

- A data frame in wide format (see an example below). If it's in wide format, you can still use `pivot_wider()` from the `tidyr` package to make wider.
- A date variable in `yyyy-mm-dd` format that will plotted in the x-axis.

Let's work with the following data frame (shortened) provided by [Mike Bostock in his stacked area chart example](https://observablehq.com/@d3/stacked-area-chart):

```{r}
data <- data.frame(
  date = c(
    "2000-01-01", "2000-02-01", "2000-03-01", "2000-04-01",
    "2000-05-01", "2000-06-01", "2000-07-01",
    "2000-08-01", "2000-09-01", "2000-10-01"
  ),
  Trade = c(
    2000,1023, 983, 2793, 1821, 1837, 1792, 1853, 791, 739
  ),
  Manufacturing = c(
    734, 694, 739, 736, 685, 621, 708, 685, 667, 693
  ),
  Leisure = c(
    1782, 1779, 1789, 658, 675, 833, 786, 675, 636, 691
  ),
  Agriculture = c(
    655, 587,623, 517, 561, 2545, 636, 584, 559, 2504
  )
)

data
```


Note that when running `stackedAreaChart()` all the variables available within the considered data frame will be plotted. If you want to restrict the plotting to only specific variables, just drop the unneeded columns:


```{r}
stackedAreaChart(
  data = data,
  x = "date",
  legendTextSize = 14
  )
```

You can modify the color scheme using the `colorCategory` parameter:


```{r}
stackedAreaChart(
  data = data,
  x = "date",
  legendTextSize = 14,
  curve = "curveCardinal",
  colorCategory = "Accent",
  bgcol = "white",
  stroke = "black",
  strokeWidth = 1
  )
```


```{r}
stackedAreaChart(
  data = data,
  x = "date",
  legendTextSize = 14,
  curve = "curveBasis",
  colorCategory = "Set3",
  bgcol = "black",
  axisCol = "white",
  xticks = 4,
  stroke = "black"
  )
```

You can find list of D3 categorical color schemes [here](https://github.com/d3/d3-scale-chromatic#categorical)

Finally, if you hover over the chart you'll notice a tooltip that identified the different area categories.

# `barChartRace()`

This function allows you to create an animated bar chart race. `barChartRace()` is similar to `barChart()` but takes a third variable mapped to the time dimension, with options for styling transitions.

Let's make a bar chart race of population growth among various countries using a subset of the `gapminder` dataset from the [{gapminder} package](https://github.com/jennybc/gapminder):

```{r, eval = FALSE}
gapminder_subset <- gapminder::gapminder %>%
  select(country, year, pop) %>% 
  filter(country %in% c("Japan", "Mexico", "Germany", "Brazil", "Philippines", "Vietnam")) %>%
  mutate(pop = pop/1e6)


gapminder_subset %>%
  slice_sample(n = 10)

#>    year       pop     country
#> 1  2007  91.07729 Philippines
#> 2  1997  76.04900     Vietnam
#> 3  1972 107.18827       Japan
#> 4  1967  39.46391     Vietnam
#> 5  1952  30.14432      Mexico
#> 6  1987 142.93808      Brazil
#> 7  1997 168.54672      Brazil
#> 8  1962  41.12148      Mexico
#> 9  1952  69.14595     Germany
#> 10 1957  91.56301       Japan
```


```{r, echo = FALSE}
gapminder_subset <- data.frame(
  year = c(
    1952L,1957L,1962L,1967L,1972L,1977L,
    1982L,1987L,1992L,1997L,2002L,2007L,1952L,1957L,1962L,
    1967L,1972L,1977L,1982L,1987L,1992L,1997L,2002L,2007L,
    1952L,1957L,1962L,1967L,1972L,1977L,1982L,1987L,1992L,
    1997L,2002L,2007L,1952L,1957L,1962L,1967L,1972L,1977L,
    1982L,1987L,1992L,1997L,2002L,2007L,1952L,1957L,1962L,
    1967L,1972L,1977L,1982L,1987L,1992L,1997L,2002L,2007L,
    1952L,1957L,1962L,1967L,1972L,1977L,1982L,1987L,1992L,
    1997L,2002L,2007L
  ),
  pop = c(
    56.60256,65.551171,76.03939,88.049823,
    100.840058,114.313951,128.962939,142.938076,155.975974,
    168.546719,179.914212,190.010647,69.145952,71.019069,73.739117,
    76.368453,78.717088,78.160773,78.335266,77.718298,
    80.597764,82.011073,82.350671,82.400996,86.459025,91.563009,
    95.831757,100.825279,107.188273,113.872473,118.454974,
    122.091325,124.329269,125.956499,127.065841,127.467972,30.144317,
    35.015548,41.121485,47.995559,55.984294,63.759976,
    71.640904,80.122492,88.11103,95.895146,102.479927,108.700891,
    22.438691,26.072194,30.325264,35.3566,40.850141,46.850962,
    53.456774,60.017788,67.185766,75.012988,82.995088,91.077287,
    26.246839,28.998543,33.79614,39.46391,44.655014,50.533506,
    56.142181,62.826491,69.940728,76.048996,80.908147,
    85.262356
  ),
  country = as.factor(c(
    "Brazil","Brazil",
    "Brazil","Brazil","Brazil","Brazil","Brazil",
    "Brazil","Brazil","Brazil","Brazil","Brazil","Germany",
    "Germany","Germany","Germany","Germany",
    "Germany","Germany","Germany","Germany","Germany",
    "Germany","Germany","Japan","Japan","Japan","Japan",
    "Japan","Japan","Japan","Japan","Japan","Japan",
    "Japan","Japan","Mexico","Mexico","Mexico",
    "Mexico","Mexico","Mexico","Mexico","Mexico",
    "Mexico","Mexico","Mexico","Mexico","Philippines",
    "Philippines","Philippines","Philippines","Philippines",
    "Philippines","Philippines","Philippines",
    "Philippines","Philippines","Philippines","Philippines",
    "Vietnam","Vietnam","Vietnam","Vietnam",
    "Vietnam","Vietnam","Vietnam","Vietnam","Vietnam",
    "Vietnam","Vietnam","Vietnam"
  ))
)
```

In this example, we simply pass call `barChartRace()` like `barChart()`, but with an additional variable mapped to the time dimension specified with `time = year`:

```{r}
gapminder_subset %>%
  barChartRace(
    x = "pop",
    y = "country",
    time = "year",
    ytitle = "Country",
    xtitle = "Population (in millions)",
    title = "Bar chart race of country populations"
  )
```

You can also stylize transitions with the `frameDur`, `transitionDur`,  and `ease` arguments. For example, setting the time spent pausing on each frame to zero with `frameDur = 0` will create a smooth animation:

```{r}
gapminder_subset %>%
  barChartRace(
    x = "pop",
    y = "country",
    time = "year",
    transitionDur = 1000,
    frameDur = 0,
    ytitle = "Country",
    xtitle = "Population (in millions)",
    title = "Bar chart race of country populations"
  )
```

As you might have noticed, the value of the column passed to the `time` argument is automatically labelled at the bottom-right corner of the plot panel. We can stylize this with a list of options passed to the `timeLabelOpts` argument (or turn it off with `timeLabel = FALSE`). We also give the bars a little bounce here with `ease = "BackInOut"` for fun.

```{r}
gapminder_subset %>%
  barChartRace(
    x = "pop",
    y = "country",
    time = "year",
    ease = "BackInOut",
    ytitle = "Country",
    xtitle = "Population (in millions)",
    title = "Bar chart race of country populations",
    timeLabelOpts = list(
      size = 40,
      prefix = "Year: ",
      xOffset = 0.2
    )
  )
```


# Using `ddplot` in `Shiny` 

If you want to add reactivity to your `ddplot` plots, use the `uiOutput` and `renderUI` in order to render a frame, note that you'll need to have a `www` folder, below an example: 


```{r, eval=FALSE}
library(shiny)
library(ddplot)
library(r2d3)

ui <- fluidPage(

  shiny::h2("Example ddplot application"),

  shiny::selectInput(
    inputId = "colors",
    label = NULL,
    choices = colors(),
    selected = "springgreen3"
  ),

  shiny::br(),

  shiny::sliderInput(
    inputId = "slider",
    label = NULL,
    min = 1,
    max = 10,
    value = 3
  ),
  shiny::br(),
  mainPanel(
    uiOutput("ddplot_ui")
  )
)

server <- function(input, output) {
  output$ddplot_ui <- renderUI({
    widget <- ddplot::scatterPlot(
      data = iris,
      x = "Sepal.Length",
      y = "Sepal.Width",
      size = input$slider,
      col = input$colors
    )
    htmlwidgets::saveWidget(widget, "www/temp_ddplot.html", selfcontained = TRUE)
    tags$iframe(src = "temp_ddplot.html", width = "100%", height = "400px", frameborder = 0)
  })
}

shinyApp(ui = ui, server = server)
```










