---
title: "Start"
output: html_document
vignette: >
  %\VignetteIndexEntry{Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ddplot)
```

`D3.js` is a famous JavaScript library that allows one to create extremely flexible SVG graphics however `D3` has (at least according to me) a pretty steep learning curve. Further, in order to understand some core concepts, one need to have some basics in `HTML`, `CSS` and `JavaScript`. `ddplot` aims to simply the process using a set of functions that render several graphics using a simple `R` API. Finally, `ddplot` is built upon the amazing `r2d3` package which makes it a breeze to interface `D3.js` with `R`, so a big thanks to the developers.  

# `scatterPlot()`

Let's work with the `mpg` data frame from the `ggplot2` package. 

```{r fig.align='center', message=FALSE, warning=FALSE}
library(ggplot2) # needed for the mpg data frame

scatterPlot(
  data = mpg, 
  x = "hwy", 
  y = "cty", 
  xtitle = "hwy variable", 
  ytitle = "cty variable", 
  title = "cty and hwy relationship", 
  titleFontSize = 20
)

```

In comparison to `ggplot2`, graphics' customization in `ddplot` is limited nonetheless you get a fully vectorized SVG which is cool. 



```{r, fig.align='center'}
scatterPlot(
  data = mpg, 
  x = "displ", 
  y = "cty", 
  col = "tomato", 
  bgcol = "pink", 
  size = 3, 
  stroke = "royalblue", 
  strokeWidth = 1,
  xtitle = "displ variable", 
  ytitle = "cty variable", 
  xticks = 3, 
  yticks = 3)

```

# `barChart()`

The `barChat()` function allows you to create bar charts however you need to make the aggregation beforehand. In the following example, we will plot the average `cty` for each `manufacturer` using the `dplyr` package.   

```{r fig.align='center', message=FALSE, warning=FALSE}
library(dplyr)

mpg %>% group_by(manufacturer) %>% 
  summarise(mean_cty = mean(cty)) %>% 
  barChart(
    x = "manufacturer", 
    y = "mean_cty", 
    xFontSize = 10,
    yFontSize = 10,
    fill = "orange", 
    strokeWidth = 2, 
    ytitle = "average cty value", 
    title = "Average City Miles per Gallon by manufacturer"
  )
```

The bars can be easily sorted in `ascending` or `descending` order using the `sort` parameter: 


```{r message=FALSE, warning=FALSE}
mpg %>% group_by(manufacturer) %>% 
  summarise(mean_cty = mean(cty)) %>% 
  barChart(
    x = "manufacturer", 
    y = "mean_cty", 
    sort = "ascending",
    xFontSize = 10, 
    yFontSize = 10,
    fill = "orange", 
    strokeWidth = 1, 
    ytitle = "average cty value", 
    title = "Average City Miles per Gallon by manufacturer", 
    titleFontSize = 16
  )
```


# `horzBarChart()` 

If you've many categories, it might be a good idea to go for a horizontal bar chart. It has the same parameters as the `barChart()` function except that the x-axis parameter is named `value` and the y-axis parameter named `label`, this naming convention aims to mitigate some confusion that can arise. Also, I've made the choice to not include a title in the label axis as I think it's cumbersome and the labels can tell the story. 

If we want to replicate the above graphic in a horizontal way, we can do: 

```{r}
mpg %>% group_by(manufacturer) %>% 
  summarise(mean_cty = mean(cty)) %>% 
  horzBarChart(
    label = "manufacturer", 
    value = "mean_cty", 
    sort = "ascending",
    labelFontSize  = 10, 
    valueFontSize = 10,
    fill = "orange", 
    stroke = "crimson",
    strokeWidth = 1, 
    valueTitle  = "average cty value", 
    title = "Average City Miles per Gallon by manufacturer", 
    titleFontSize = 16
  )
```

As in `barChart()`, we can aslo sort in descending order: 


```{r}
mpg %>% group_by(manufacturer) %>% 
  summarise(mean_cty = mean(cty)) %>% 
  horzBarChart(
    label = "manufacturer", 
    value = "mean_cty", 
    sort = "descending",
    labelFontSize  = 10, 
    valueFontSize = 10,
    bgcol = "white",
    fill = "royalblue", 
    stroke = "white",
    strokeWidth = 1, 
    valueTitle  = "average cty value", 
    title = "Average City Miles per Gallon by manufacturer", 
    titleFontSize = 16
  )
```


# `lollipopChart()` 

lollipop chart follows the same behavior as bar charts but instead of bars you get lollipops, hence the name. Below an example of a lollipop chart with `ddplot`:  


```{r}

mpg %>% group_by(drv) %>% 
  summarise(median_cty = median(cty)) %>% 
  lollipopChart(
    x = "drv", 
    y = "median_cty", 
    sort = "ascending", 
    xtitle = "drv variable", 
    ytitle = "median cty", 
    title = "Median cty per drv", 
    xFontSize = 20
  )

```


It's possible to grasp the distribution of some variable according to a specific categorical variable using the same function: 


```{r}

mpg %>% filter(year == 2008) %>%   
lollipopChart(
    x = "manufacturer", 
    y = "hwy", 
    circleFill = 'red', 
    circleStroke = 'orange',
    circleRadius = 5, 
    sort = "none", 
    xFontSize = 10
  )

```

From above, it's quite easy to notice that although Toyota has two cars with high highway miles per galon (hwy), it also produces many other vehicles with poor hwy. 


# `horzLollipop()` 

Same with bar charts, if you have a variable that has many categorical values, you can work with the reversed version of `lollipopChart()` which is `horzLollipop()`: 


```{r}
mpg %>% group_by(manufacturer) %>% 
  summarise(median_cty = median(cty)) %>% 
  horzLollipop(
    label = "manufacturer", 
    value = "median_cty", 
    sort = "descending")

```
You can also do: 



```{r}

mpg %>% filter(year == 2008) %>%   
horzLollipop(
    label = "manufacturer", 
    value = "hwy", 
    circleFill = 'red', 
    circleStroke = 'orange',
    circleRadius = 5, 
    sort = "none"
  )

```


# `pieChart()` 

Pie charts and donut charts are pretty straightforward to set up. We'll use a sample from the `starwars` data frame to plot a simple pie chart. 


```{r}
# starwars is part of the dplyr data frame

mini_starwars <- starwars %>% tidyr::drop_na(mass) %>% 
  sample_n(size = 5) # getting 5 random values

pieChart(
  data = mini_starwars, 
  value = "mass",
  label = "name"
)

```

Using the `padRadius`, `padAngle` and `cornerRadius` parameters, one can get fanciers pie charts: 

```{r}
pieChart(
  data = mini_starwars, 
  value = "mass",
  label = "name", 
  padRadius = 200, 
  padAngle = 0.1, 
  cornerRadius = 50, 
  innerRadius = 10
)
```
If you need a donut chart, you just need to play with the `innerRadius` parameter: 

```{r}

pieChart(
  data = mini_starwars, 
  value = "mass",
  label = "name", 
  innerRadius = 120, 
  cornerRadius = 20, 
  title = "5 Starwars characters ranked by their mass",
  titleFontSize = 16, 
  bgcol = "yellow"
)

```

# `lineChart()` 

The `lineChart()` function is used to plot time series data. The use must provide a `date` variable that has the `yyyy-mm-dd` format. In the following example, we'll use the `Air Passenger` built-in `ts` data and convert it to a classical data frame: 


```{r}

# 1. converting AirPassengers to a tidy data frame

airpassengers <- data.frame(
  passengers = as.matrix(AirPassengers),
  date= zoo::as.Date(time(AirPassengers))
)

# 2. plotting the line chart
lineChart(
  data = airpassengers, 
  x = "date", 
  y = "passengers"
)

```

# `areaChart()` 

`areaChart()` works similarly except that instead of a line you get an area. 


```{r}

# 1. converting AirPassengers to a tidy data frame

airpassengers <- data.frame(
  passengers = as.matrix(AirPassengers),
  date= zoo::as.Date(time(AirPassengers))
)

# 2. plotting the area chart
areaChart(
  data = airpassengers, 
  x = "date", 
  y = "passengers", 
  fill = "purple", 
  bgcol = "white"
)

```

# `areaBand()` 

`areaBand()` lets you plot a filled area between two y-values. For the sake of the example, let's create an additional column `passengers_upper` that has an additional 40 passengers for each observation: 

```{r}

airpassengers <- data.frame(
  passengers_lower = as.matrix(AirPassengers),
  passengers_upper = as.matrix(AirPassengers) + 40,
  date= zoo::as.Date(time(AirPassengers))
)

areaBand(
  data = airpassengers,
  x = "date",
  yLower = "passengers_lower", 
  yUpper = "passengers_upper", 
  fill = "yellow", 
  stroke = "black"
)



```

# More to Come ... 

